---
title: "ECON4200 - Hong Kong Statutory Minimum Wage Analysis\nEducation"
output:
  pdf_document:
    latex_engine: xelatex
    fig_crop: false

header-includes:
  - \usepackage[table]{xcolor}
  - \usepackage{booktabs}
  - \usepackage{caption}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{colortbl}

date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  fig.align = "center",
  dev = "cairo_pdf",
  out.width = "100%",
  cache = FALSE,
  fig.path = "figs/"
)
options(scipen = 999)
```



# setting up again
## Load Libraries & Define Mappings
load lib
```{r}
library(tidyverse)
library(lubridate)
library(ggridges)
library(Hmisc)
library(kableExtra)
library(knitr)
```
done load lib



load earning map
```{r, echo=TRUE}
earnings_map <- tibble(
  code = str_pad(1:40, width = 2, pad = "0"),
  midpoint_hkd = c(1000,2500,3500,4500,5500,6500,7500,8500,9500,
                   10500,11500,12500,13500,14500,15500,16500,17500,
                   18500,19500,21000,23000,25000,27000,29000,31000,
                   33000,35000,37000,39000,41000,43000,45000,47000,
                   49000,55000,65000,75000,85000,95000,120000)
)

other_earnings_map <- tibble(
  code = str_pad(1:11, width = 2, pad = "0"),
  midpoint_hkd = c(500,1500,2500,3500,4500,5500,6500,7500,8500,9500,12000)
)
```
done load earning map


## Data Loading & Processing
Load all files + calculate est hourly wage
same method used
```{r}
files <- list.files("data_csv", pattern = "*.csv", full.names = TRUE)
df_raw <- files %>% map_df(~read_csv(.x, col_types = cols(.default = "c")))

df <- df_raw %>%
  mutate(
    year_quarter = str_trim(as.character(year_quarter)),
    year = as.integer(str_sub(year_quarter, 1, 4)),
    quarter = as.integer(str_sub(year_quarter, 5, 5)),
    date = yq(paste0(year, "-Q", quarter)),
    
    across(c(earnings_main, earnings_employed, earnings_other),
           ~ str_pad(as.character(.), width = 2, pad = "0", side = "left")),
    
    earnings_main_hkd     = earnings_map$midpoint_hkd[match(earnings_main, earnings_map$code)],
    earnings_employed_hkd = earnings_map$midpoint_hkd[match(earnings_employed, earnings_map$code)],
    earnings_other_hkd    = other_earnings_map$midpoint_hkd[match(earnings_other, other_earnings_map$code)],
    
    across(c(hours_main_employment, hours_7days),
           ~ case_when(.x %in% c("999","99","","NA") ~ NA_real_, TRUE ~ as.numeric(.x))),
    
    hourly_wage = case_when(
      !is.na(earnings_main_hkd) & !is.na(hours_main_employment) & hours_main_employment > 0 ~
        earnings_main_hkd / (hours_main_employment * 4.345),
      !is.na(earnings_employed_hkd) & !is.na(hours_7days) & hours_7days > 0 ~
        earnings_employed_hkd / (hours_7days * 4.345),
      TRUE ~ NA_real_
    ),
    
    is_employed       = economic_status %in% c("1","2","3"),
    working_age       = age %in% str_pad(4:13, 2, pad = "0"),
    not_foreign_helper = foreign_helper != "1" | is.na(foreign_helper),
    not_underemployed  = underemployed != "1",
    weight = as.numeric(grossing_up_factor) / 1e12
  ) %>%
  filter(!is.na(date), year >= 1985)


```
done

ranges mapping
```{r}
df_focus <- df %>%
  filter(is_employed, working_age, not_foreign_helper, not_underemployed,
         !is.na(hourly_wage), hourly_wage > 0,
         hours_main_employment >= 8 | is.na(hours_main_employment)) %>%
  mutate(
    year_group = case_when(
      year <= 1989 ~ "1985-1989",
      year >= 1990 & year <= 1994 ~ "1990-1994",
      year >= 1995 & year <= 1999 ~ "1995-1999",
      year >= 2000 & year <= 2004 ~ "2000-2004",
      year >= 2005 & year <= 2009 ~ "2005-2009",
      year >= 2010 & year <= 2012 ~ "2010-2012",
      TRUE ~ "Other"
    )
  )
```
done



## CPI Processing (Real Wages)
```{r}
raw_cpi <- read_csv("Table 510-60001_en.csv", skip = 4)

cleaned_cpi <- raw_cpi %>%
  rename(Year = 1, Month = 2, Comp_Index = 3, CPIA_Index = 6) %>%
  fill(Year) %>%
  filter(!is.na(Month)) %>%
  mutate(Year = as.numeric(Year)) %>% 
  mutate(across(c(Comp_Index, CPIA_Index), as.numeric)) %>%
  filter(Year >= 1985 & Year <= 2012)

cpi_quarterly <- cleaned_cpi %>%
  mutate(quarter = case_when(
    Month %in% c("Jan", "Feb", "Mar") ~ 1,
    Month %in% c("Apr", "May", "Jun") ~ 2,
    Month %in% c("Jul", "Aug", "Sep") ~ 3,
    Month %in% c("Oct", "Nov", "Dec") ~ 4
  )) %>%
  dplyr::group_by(Year, quarter) %>%
  dplyr::summarise(
    cpi_comp = mean(Comp_Index, na.rm = TRUE),
    cpi_a = mean(CPIA_Index, na.rm = TRUE),
    .groups = "drop"
  )

ref_cpi <- cpi_quarterly %>%
  filter(Year == 2011, quarter == 2) %>%
  pull(cpi_comp)

print(paste("2011 Q2 Reference CPI Value is:", ref_cpi))
```
done

## Add Real Wage to Main Dataset
```{r}
df_focus <- df_focus %>%
  left_join(cpi_quarterly, by = c("year" = "Year", "quarter" = "quarter")) %>%
  mutate(real_hourly_wage = (hourly_wage / cpi_comp) * ref_cpi)
```
done adding real wage

## Testing to reconfirm is same as previous rmd file
```{r data-validation}
cat("=== Basic Data Overview ===\n")
cat("Total rows in raw data:", nrow(df_raw), "\n")
cat("Total rows after processing:", nrow(df), "\n")
cat("Rows in focus sample:", nrow(df_focus), "\n\n")

cat("Years covered:\n")
print(df_focus %>% count(year, quarter) %>% arrange(year, quarter))

cat("\n=== Key Variable Summary ===\n")
summary(df_focus %>% select(hourly_wage, real_hourly_wage, weight, 
                            hours_main_employment, earnings_main_hkd))

cat("\n=== Missing Values (%) ===\n")
df_focus %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing_%") %>%
  filter(`Missing_%` > 0) %>%
  arrange(desc(`Missing_%`)) %>%
  print()

cat("\n=== Hourly Wage Distribution (Nominal) ===\n")
df_focus %>%
  summarise(
    Min = min(hourly_wage, na.rm = TRUE),
    P5 = quantile(hourly_wage, 0.05, na.rm = TRUE),
    Median = median(hourly_wage, na.rm = TRUE),
    Mean = mean(hourly_wage, na.rm = TRUE),
    P95 = quantile(hourly_wage, 0.95, na.rm = TRUE),
    Max = max(hourly_wage, na.rm = TRUE),
    N = n()
  ) %>% 
  kable(digits = 2)
```

### 1. Check Earnings Mapping
```{r check-earnings-mapping}
cat("Unique earnings_main codes and their mapped values:\n")
df_focus %>%
  count(earnings_main, earnings_main_hkd) %>%
  arrange(earnings_main) %>%
  print(n = 30)
```

### 2. Check Weighting Factor
```{r check-weights}
cat("Total weighted population (should be around Hong Kong's employed population):\n")
sum(df_focus$weight, na.rm = TRUE) %>% round(3)
```


### 3. Sample of Raw vs Processed Data
yea this one isnt going to be same but uh it exist
```{r sample-data}
df_focus %>%
  select(year, quarter, earnings_main, earnings_main_hkd, 
         hours_main_employment, hourly_wage, real_hourly_wage) %>%
  slice_sample(n = 10) %>%
  kable(digits = 2)
```



Now actual content

Cont.
The aggregate data shows a continued improvement in wages and a reduction in the low-wage share around 2011, the effect appears modest in the overall population. This raises the question of whether the minimum wage had a stronger impact on specific vulnerable groups.
Therefore, the next step is to conduct a more disaggregated analysis by examining subgroups such as:
- Workers with lower education levels

# Education Grouping (Handle code change in 2012)

From Q1 1985 to Q4 2011
Code Description
1 No schooling / pre-primary
2 Primary
3 Lower secondary
4 Upper secondary
5 Six form
6 Technical / vocational training
7 Tertiary (non-degree courses)
8 Tertiary (First degree courses)
9 Tertiary (Post-degree courses)

From Q1 2012 to current
Code Description
1 No schooling / pre-primary
2 Primary
3 Lower secondary
4 Upper secondary
5 Post-secondary – diploma /
certificate including craft
6 Post-secondary – sub-degree
7 Post-secondary – first degree
8 Post-secondary – post-degree


Then we have to do some recode mapping from numbers to words 

Codes 1–4: No recoding needed, meanings identical across both periods

Code 5 (pre-2012): Kept as "Sixth Form", this is academic pre-university, distinct from upper secondary (code 4)

Code 5 (post-2012): Labeled "Post-secondary Diploma/Certificate", different meaning from pre-2012 code 5, so kept separate

Code 6 (pre-2012): Labeled "Technical / Vocational", preserved as distinct category

Code 6 (post-2012): Labeled "Post-secondary Sub-degree", different meaning from pre-2012 code 6, kept separate

Code 7 (pre-2012): Labeled "Tertiary Non-degree", below bachelor level, preserved as distinct category

Code 7 (post-2012): Mapped to "Bachelor's Degree", conceptually equivalent to pre-2012 code 8

Code 8 (pre-2012): Mapped to "Bachelor's Degree", conceptually equivalent to post-2012 code 7

Code 8 (post-2012): Mapped to "Postgraduate", conceptually equivalent to pre-2012 code 9

Code 9 (pre-2012): Mapped to "Postgraduate", no equivalent in post-2012 scheme

Codes 99, 999, blank, NA: Coded as "Unknown / Missing"








```{r education-grouping}
# Education recoding with clearer naming
df_focus <- df_focus %>%
  mutate(
    education_code = as.character(education),
    
    educ_group = case_when(
      # Primary and below
      education_code %in% c("1", "2") ~ "No Schooling / Primary",
      education_code == "3"           ~ "Lower Secondary",
      education_code == "4"           ~ "Upper Secondary",
      
      # Code 5 transition
      education_code == "5" & year <= 2011 ~ "Sixth Form",
      education_code == "5" & year >= 2012 ~ "Post-secondary Diploma/Certificate",
      
      # Code 6 transition
      education_code == "6" & year <= 2011 ~ "Technical / Vocational",
      education_code == "6" & year >= 2012 ~ "Post-secondary Sub-degree",
      
      # Code 7 transition
      education_code == "7" & year <= 2011 ~ "Tertiary Non-degree",
      education_code == "7" & year >= 2012 ~ "First Degree",
      
      # Code 8 transition
      education_code == "8" & year <= 2011 ~ "First Degree",
      education_code == "8" & year >= 2012 ~ "Post Degree",
      
      # Code 9 (pre-2012 only)
      education_code == "9" & year <= 2011 ~ "Post Degree",
      
      # Missing
      is.na(education) | education_code %in% c("", "99", "999") ~ "Unknown / Missing",
      TRUE ~ "Other"
    ),
    
    educ_broad = case_when(
      # Lower education
      education_code %in% c("1", "2", "3") ~ "Lower Education",
      
      # Secondary + Vocational (including non-degree tertiary pre-2012)
      education_code %in% c("4", "5", "6") ~ "Secondary + Vocational",
      (education_code == "7" & year <= 2011) ~ "Secondary + Vocational",
      
      # Tertiary (Degree) - First Degree and Post Degree
      (education_code == "7" & year >= 2012) ~ "Tertiary (Degree)",
      (education_code == "8" & year <= 2011) ~ "Tertiary (Degree)",
      (education_code == "8" & year >= 2012) ~ "Tertiary (Degree)",
      (education_code == "9" & year <= 2011) ~ "Tertiary (Degree)",
      
      TRUE ~ "Unknown / Missing"
    )
  )
```

# Education distribution

Let's check the education distribution:

```{r}
# Check education distribution
df_focus %>%
  count(educ_group, educ_broad) %>%
  arrange(desc(n)) %>%
  mutate(pct = round(n / sum(n) * 100, 2))
```

```{r}
cat("Percentage of Unknown education in focus sample:", 
    round(mean(df_focus$educ_group == "Unknown / Missing") * 100, 2), "%\n")
```

2.4% unknown edu, we can filter them to make final graphs/tables cleaner

```{r}
df_focus <- df_focus  %>% filter(educ_group != "Unknown / Missing")

cat("Percentage of Unknown education in focus sample after filtering yes this is going to be 0%:", 
    round(mean(df_focus$educ_group == "Unknown / Missing") * 100, 2), "%\n")
```

# Figure 1: Proportion Below HK 28 by Education Group

```{r fig.width=15, fig.height=15, out.width="100%"}
# Proportion below real $28 by education
prop_educ <- df_focus %>%
  filter(educ_group != "Unknown / Missing") %>%
  group_by(year, educ_group) %>%
  summarise(
    below_28 = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  ) %>%
  filter(year >= 2002)

ggplot(prop_educ, aes(x = year, y = below_28, color = educ_group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.2) +
  geom_text(aes(label = round(below_28, 1)), 
            vjust = -0.8, size = 3.5, show.legend = FALSE) +   # ← Data labels
  scale_x_continuous(breaks = seq(2002, 2012, by = 2)) +
  labs(title = "Proportion Earning Below Real HK$28 by Education Level",
       subtitle = "Real wages in 2011 prices",
       x = "Year", y = "Proportion (%)") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```



```{r fig.width=15, fig.height=15, out.width="100%"}
# Identify education groups with sufficient data points
valid_groups <- prop_educ %>%
  group_by(educ_group) %>%
  summarise(n_years = n(), .groups = "drop") %>%
  filter(n_years >= 3) %>%
  pull(educ_group)

# Filter and plot
prop_educ_filtered <- prop_educ %>%
  filter(educ_group %in% valid_groups)

ggplot(prop_educ_filtered, aes(x = year, y = below_28, color = educ_group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.2) +
  geom_text(aes(label = round(below_28, 1)), 
            vjust = -0.8, size = 3.5, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(2002, 2012, by = 2)) +
  labs(title = "Proportion Earning Below Real HK$28 by Education Level",
       subtitle = "Real wages in 2011 prices | Groups with sufficient pre/post data",
       x = "Year", y = "Proportion (%)") +
  
  geom_vline(xintercept = 2011, linetype = "dashed", color = "red", linewidth = 0.8) +
annotate("text", x = 2011, y = Inf, label = "SMW", color = "red", size = 4, hjust = -0.2, vjust = 1.5)+
  
  
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```
## Key Findings from the Education Analysis

Yes the changes are quite significant to the point you can directly use your eyes to see some big different but anyways

### 1. Strongest Impact on the Least Educated Workers

Workers with No Schooling / Primary education consistently had the highest share of low-wage earners in their own group, like maybe no schooling 100 people has like 25 people are wage < 28 hkd vs upper school maybe 100 people has like 5 people only are wage < 28 hkd etc

BUT the thing is, in 2011 compared to 2010 there is a big drop of proportion, like 25/100 ppl then 2011 become 17/100 which is quite some huge improvement

### 2. Limited Impact on Better-Educated Workers

I know i know the lines are "broke off" in 2012 but ignoring the broken line you can kinda know which is which

but either way that's not the most important, the most important is that

Upper Secondary, Sixth Form, and Post-secondary groups had much lower exposure to wages below $28 even before the policy.

Tertiary-educated workers (Bachelor’s, Postgraduate, etc.) had very low proportions (< 3–4%) throughout the period, with almost no meaningful change after 2011.

### 3. The Policy Accelerated an Existing Trend

There was already a gradual downward trend in low-wage employment before 2011 (due to economic growth and structural changes) yes i know there are highs and ups and downs but the overall trend is decreasing.

However, the speed of decline increased noticeably in 2010 vs 2011, especially for No Schooling/Primary and Lower Secondary groups, consistent with the minimum wage taking effect in May 2011.




# YoY % Change Tables and actual values by Education Group

Let's see the YoY and actual of different edu group yes it is going to be % change = (new - old)/old x 100%

we see actual values cuz the graphs even i fig.width=10, fig.height=15, out.width="100%" still has some data label overlapping each other

yes if they are some newly changed edu group in 2012 data then there is NOT going to be previous year comparison so they maybe NA on the % change, same reason if you are seeing removed and 2012 is gone

but we mainly care about the no schooling / primary AND lower secondary cuz 1. they are low edu 2. their % change are largest


```{r}

educ_groups <- df_focus %>%
  filter(educ_group != "Unknown / Missing") %>%
  pull(educ_group) %>% unique()

for (group in sort(educ_groups)) {
  cat("\n##", group, "\n\n")
  
  table_data <- df_focus %>%
    filter(educ_group == group, year >= 2007, year <= 2012) %>%   # ← Include 2007
    group_by(year) %>%
    summarise(
      `Below $28` = round(weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100, 1),
      `Below $35` = round(weighted.mean(real_hourly_wage < 35, w = weight, na.rm = TRUE) * 100, 1),
      `Below $40` = round(weighted.mean(real_hourly_wage < 40, w = weight, na.rm = TRUE) * 100, 1),
      .groups = "drop"
    ) %>%
    arrange(year) %>%
    mutate(
      `YoY $28` = round((`Below $28` - lag(`Below $28`)) / lag(`Below $28`) * 100, 1),
      `YoY $35` = round((`Below $35` - lag(`Below $35`)) / lag(`Below $35`) * 100, 1),
      `YoY $40` = round((`Below $40` - lag(`Below $40`)) / lag(`Below $40`) * 100, 1)
    ) %>%
    filter(year >= 2008) %>%   # Show only 2008–2012
    select(year, 
           `Below $28`, `YoY $28`, 
           `Below $35`, `YoY $35`, 
           `Below $40`, `YoY $40`)
  
  print(kable(table_data, 
              caption = paste("Proportion Below Thresholds & YoY % Change -", group),
              align = "r"))
  
  cat("\n")
}
```


# Before vs After May 2011: Changes by Educational Attainment

To more directly assess the immediate impact of the Statutory Minimum Wage implemented in May 2011, we compare the proportion of workers earning below various hourly wage thresholds in the period **2008Q1–2011Q1 (Before)** VS **2011Q2–2012Q4 (After)**.



Why We Show 3 Thresholds ($28, $35, $40) Instead of Just $28

1. To show the depth of the policy's impact

$28 = The actual Statutory Minimum Wage (the legal floor).

$35 and $40 = Slightly higher thresholds to see if the policy also helped workers who were earning just above the minimum wage (a "ripple effect" or "spillover effect").

If only the $28 bar drops significantly while $35 and $40 stay almost the same, it means the minimum wage mainly pushed people up to the floor, but didn't help much beyond that.
If all three drop, it suggests a broader compression of the lower wage distribution.


2. To make the analysis more robust and convincing

Relying on only one threshold ($28) is risky — it could be affected by measurement error or specific data quirks.

Showing multiple thresholds makes your conclusion stronger and more credible to readers/professors.

3. Academic / Policy Relevance

In minimum wage research, it's very common to examine effects at:

The exact minimum wage
10–20% above the minimum wage


This helps answer:
→ "Did the policy only affect the very bottom, or did it lift a wider group of low-wage workers?"





**NO need to keep all three separate plots in the final report if space is limited.**



## Below $28 graph
MAKE IT 2011Q1 VS 2011Q3

```{r}
# Before vs After by Education - Multiple Thresholds (Vertical)
event_educ_multi <- df_focus %>%
  filter(year >= 2008, year <= 2012,
         educ_group != "Unknown / Missing") %>%
  mutate(
    period = if_else(date < as.Date("2011-05-01"), 
                     "Before (Real)", 
                     "After (Real)")
  ) %>%
  group_by(period, educ_group) %>%
  summarise(
    `Below $28` = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100,
    `Below $35` = weighted.mean(real_hourly_wage < 35, w = weight, na.rm = TRUE) * 100,
    `Below $40` = weighted.mean(real_hourly_wage < 40, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  )

event_educ_long <- event_educ_multi %>%
  pivot_longer(cols = starts_with("Below"), 
               names_to = "threshold", 
               values_to = "proportion")

# Plot 1: Below $28
ggplot(event_educ_long %>% filter(threshold == "Below $28"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$28",
       subtitle = "Before vs After May 2011",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (Real)" = "grey60", "After (Real)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



## Below $35 graph

```{r}
# Plot 2: Below $35
ggplot(event_educ_long %>% filter(threshold == "Below $35"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$35",
       subtitle = "Before vs After May 2011",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (Real)" = "grey60", "After (Real)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Below $40 graph

```{r}
# Plot 3: Below $40
ggplot(event_educ_long %>% filter(threshold == "Below $40"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$40",
       subtitle = "Before vs After May 2011",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (Real)" = "grey60", "After (Real)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Summary Table: Before vs After by Education (Below $28)

Yes it is EXTREMELY NORMAL THE RELATIVE CHANGE IS NOT EQUAL TO PREVIOUS YOY

Relative Change (%) Compares the average of the whole Before period (2008Q1–2011Q1) vs the whole After period (2011Q2–2012Q4). This is the overall policy impact.

YoY Tables shows year-to-year changes (e.g. 2010 → 2011, 2011 → 2012). These include normal economic fluctuations

```{r}
# Fixed Summary Table
summary_table <- df_focus %>%
  filter(year >= 2008, year <= 2012,
         educ_group != "Unknown / Missing") %>%
  mutate(period = if_else(date < as.Date("2011-05-01"), "Before", "After")) %>%
  group_by(educ_group, period) %>%
  summarise(
    `Below $28 (%)` = round(weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100, 1),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = period, values_from = `Below $28 (%)`) %>%
  mutate(
    `Change (pp)` = round(After - Before, 1),
    `Relative Change (%)` = round((After - Before) / Before * 100, 1)
  ) %>%
  arrange(`Change (pp)`)

kable(summary_table, 
      caption = "Before vs After May 2011: Proportion Below Real HK dollar 28 by Education Group",
      digits = 1,
      align = "lrrrr") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

The -45.1% in the summary table is more important for the conclusion because it shows the total effect of the policy.


# RIGHT Before and After, 2011Q1 vs 2011Q3
```{r}
# === RIGHT BEFORE vs RIGHT AFTER (2011Q1 vs 2011Q3) ===

event_educ_sharp <- df_focus %>%
  filter(educ_group != "Unknown / Missing",
         (year == 2011 & quarter == 1) | (year == 2011 & quarter == 3)) %>%
  mutate(
    period = if_else(quarter == 1, 
                     "Before (2011Q1)", 
                     "After (2011Q3)")
  ) %>%
  group_by(period, educ_group) %>%
  summarise(
    `Below $28` = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100,
    `Below $35` = weighted.mean(real_hourly_wage < 35, w = weight, na.rm = TRUE) * 100,
    `Below $40` = weighted.mean(real_hourly_wage < 40, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  )

event_educ_sharp_long <- event_educ_sharp %>%
  pivot_longer(cols = starts_with("Below"), 
               names_to = "threshold", 
               values_to = "proportion")

# Plot 1: Below $28
ggplot(event_educ_sharp_long %>% filter(threshold == "Below $28"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$28",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot 2: Below $35
ggplot(event_educ_sharp_long %>% filter(threshold == "Below $35"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$35",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot 3: Below $40
ggplot(event_educ_sharp_long %>% filter(threshold == "Below $40"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$40",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




```{r}
# RIGHT BEFORE and AFTER, 2011Q1 vs 2011Q3 (Excluding Post Degree)

# Create new dataframe excluding Post Degree
event_educ_sharp_no_post <- df_focus %>%
  filter(educ_group != "Unknown / Missing",
         educ_group != "Post Degree",  # Exclude Post Degree
         (year == 2011 & quarter == 1) | (year == 2011 & quarter == 3)) %>%
  mutate(
    period = if_else(quarter == 1, 
                     "Before (2011Q1)", 
                     "After (2011Q3)")
  ) %>%
  group_by(period, educ_group) %>%
  summarise(
    `Below $28` = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100,
    `Below $35` = weighted.mean(real_hourly_wage < 35, w = weight, na.rm = TRUE) * 100,
    `Below $40` = weighted.mean(real_hourly_wage < 40, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  )

event_educ_sharp_long_no_post <- event_educ_sharp_no_post %>%
  pivot_longer(cols = starts_with("Below"), 
               names_to = "threshold", 
               values_to = "proportion")

# Plot 1: Below $28
ggplot(event_educ_sharp_long_no_post %>% filter(threshold == "Below $28"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$28 (Excluding Post Degree)",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot 2: Below $35
ggplot(event_educ_sharp_long_no_post %>% filter(threshold == "Below $35"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$35 (Excluding Post Degree)",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot 3: Below $40
ggplot(event_educ_sharp_long_no_post %>% filter(threshold == "Below $40"), 
       aes(x = educ_group, y = proportion, fill = period)) +
  geom_col(position = "dodge", width = 0.75) +
  geom_text(aes(label = round(proportion, 1)), 
            position = position_dodge(width = 0.75), 
            vjust = -0.4, size = 4.5, fontface = "bold") +
  labs(title = "Below Real HK$40 (Excluding Post Degree)",
       subtitle = "Right Before (2011Q1) vs Right After (2011Q3)",
       x = "", y = "Proportion (%)") +
  scale_fill_manual(values = c("Before (2011Q1)" = "grey60", "After (2011Q3)" = "darkblue")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



## Summary table

```{r}
sharp_summary_educ <- event_educ_sharp %>%
  select(period, educ_group, `Below $28`) %>%
  pivot_wider(names_from = period, values_from = `Below $28`) %>%
  mutate(
    `Change (pp)` = round(`After (2011Q3)` - `Before (2011Q1)`, 1),
    `Relative Change (%)` = round((`After (2011Q3)` - `Before (2011Q1)`) / `Before (2011Q1)` * 100, 1)
  ) %>%
  arrange(`Change (pp)`)

kable(sharp_summary_educ, 
      caption = "Right Before vs Right After (2011Q1 vs 2011Q3): Proportion Below HK dollar 28 by Education",
      digits = 1) %>%
  kable_styling(full_width = FALSE)
```


# Robustness Tests for Education

## Test 1: False Timing Robustness (Lowest Education Group)

Test whether the 2011 drop for "No Schooling / Primary" workers was historically unusual:


```{r}
placebo_edu_clean <- df_focus %>%
  filter(educ_group == "No Schooling / Primary", 
         year >= 1986, 
         year <= 2011) %>%
  group_by(year) %>%
  summarise(prop = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE)) %>%
  mutate(Drop = prop - lag(prop)) %>%
  filter(!is.na(Drop)) %>%  # Remove NA rows (first year has no lag)
  select(Year = year, Drop)

# Convert to percentage points
placebo_edu_clean <- placebo_edu_clean %>%
  mutate(Drop_pp = Drop * 100)

# Find largest pre-2011 drop
largest_pre_drop <- placebo_edu_clean %>%
  filter(Year < 2011) %>%
  summarise(
    largest_drop = min(Drop_pp),
    year_of_largest = Year[which.min(Drop_pp)]
  )

cat("\nLargest pre-2011 drop:\n")
print(largest_pre_drop)

# Plot
ggplot(placebo_edu_clean, aes(x = Year, y = Drop_pp, fill = ifelse(Year == 2011, "2011", "Robustness"))) +
  geom_col() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = round(Drop_pp, 1)), 
            vjust = ifelse(placebo_edu_clean$Drop > 0, -0.5, 1.5), 
            size = 3) +
  scale_fill_manual(values = c("Robustness" = "grey70", "2011" = "darkred")) +
  scale_x_continuous(breaks = seq(1986, 2011, by = 3)) +
  labs(title = "Robustness Test: No Schooling / Primary Workers",
       subtitle = paste0("Largest pre-2011 drop: ", 
                         abs(round(largest_pre_drop$largest_drop, 1)), 
                         " pp in ", largest_pre_drop$year_of_largest,
                         " | 2011 drop: ", 
                         abs(round(placebo_edu_clean$Drop_pp[placebo_edu_clean$Year == 2011], 1)), " pp"),
       x = "Year", 
       y = "Change in percentage points below HK$28 (YoY)\n(Negative = improvement)",
       fill = "") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
placebo_edu_clean <- df_focus %>%
  filter(educ_group == "No Schooling / Primary", 
         year >= 1990, 
         year <= 2011) %>%
  group_by(year) %>%
  summarise(prop = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE)) %>%
  mutate(Drop = prop - lag(prop)) %>%
  filter(!is.na(Drop)) %>%  # Remove NA rows (first year has no lag)
  select(Year = year, Drop)

# Convert to percentage points
placebo_edu_clean <- placebo_edu_clean %>%
  mutate(Drop_pp = Drop * 100)

# Find largest pre-2011 drop
largest_pre_drop <- placebo_edu_clean %>%
  filter(Year < 2011) %>%
  summarise(
    largest_drop = min(Drop_pp),
    year_of_largest = Year[which.min(Drop_pp)]
  )

cat("\nLargest pre-2011 drop:\n")
print(largest_pre_drop)

# Plot
ggplot(placebo_edu_clean, aes(x = Year, y = Drop_pp, fill = ifelse(Year == 2011, "2011", "Robustness"))) +
  geom_col() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = round(Drop_pp, 1)), 
            vjust = ifelse(placebo_edu_clean$Drop > 0, -0.5, 1.5), 
            size = 3) +
  scale_fill_manual(values = c("Robustness" = "grey70", "2011" = "darkred")) +
  #scale_x_continuous(breaks = seq(1986, 2011, by = 3)) +
  labs(title = "Robustness Test: Real Wage YoY% change of No Schooling / Primary Workers",

       x = "Year", 
       y = "Change in percentage points below HK$28 (YoY)",
       fill = "") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```


## Test 2: Quarterly Robustness for Lowest Education Group


```{r}
# Quarterly Robustness for No Schooling / Primary
quarterly_edu <- df_focus %>%
  filter(educ_group == "No Schooling / Primary", year >= 2008, year <= 2012) %>%
  group_by(year, quarter) %>%
  summarise(
    below_28 = round(weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100, 1),
    .groups = "drop"
  ) %>%
  mutate(year_quarter = paste0(year, "Q", quarter))

ggplot(quarterly_edu, aes(x = year_quarter, y = below_28, group = 1)) +
  geom_line(color = "blue", linewidth = 1.1) +
  geom_point(size = 3) +
  geom_text(aes(label = below_28), vjust = -0.8, size = 3) +
  geom_vline(xintercept = "2011Q2", linetype = "dashed", color = "red") +
  annotate("text", x = "2011Q2", y = max(quarterly_edu$below_28) + 2, 
           label = "SMW Started", color = "red", hjust = -0.1) +
  labs(title = "Quarterly Robustness Test: No Schooling / Primary Workers",
       subtitle = "Proportion Below HK$28 Over Time",
       x = "Year-Quarter", y = "Proportion Below HK$28 (%)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Small summary i guess?

1. Largest Impact on the Least Educated

Workers with No Schooling / Primary education experienced the most dramatic improvement. The proportion earning below HK$28 dropped sharply from 23.5% (Before) to 12.9% (After), a reduction of over 10 percentage points.

Lower Secondary workers also saw a substantial decline (14.1% → 8.0%).




2. Lower Secondary workers also saw a substantial decline (14.1% → 8.0%).

For workers with Upper Secondary, Sixth Form, and Post-secondary qualifications, the reductions were much smaller.

Tertiary-educated workers (Bachelor’s Degree and above) had very low exposure to sub-$28 wages even before the policy, so the impact was minimal (as expected).


3. Clearer Effect at Lower Thresholds

The policy’s effect is strongest at $28, still visible at $35, and weaker at $40. This pattern confirms that the minimum wage primarily lifted the lowest-paid workers rather than causing a broad upward shift across all low-wage earners.




# Summary?

The disaggregated analysis by educational attainment demonstrates that the Statutory Minimum Wage had a strongly progressive distributional impact.

The policy led to the largest reductions in the incidence of very low pay among workers with the lowest levels of education, aka those with No Schooling/Primary and Lower Secondary qualifications.

In the before-after comparison (2008Q1–2011Q1 vs 2011Q2–2012Q4), these groups experienced the most substantial declines in the proportion earning below HK$28 (real terms).

In contrast, the effect on tertiary-educated workers was negligible, as expected given their already higher wage levels.

**These findings suggest that the minimum wage was relatively successful in targeting and improving labour market outcomes for the most vulnerable segments of the Hong Kong workforce.**




# Limitations

While the before-after comparison and disaggregated analysis by education show a clear reduction in low-wage employment coinciding with the implementation of the minimum wage, several limitations should be noted.

First, the Labour Force Survey is a repeated cross-sectional dataset, meaning we cannot track the same individuals over time.

Second, the period 2011–2012 also coincided with economic recovery, which may have independently contributed to wage growth.

Third, because the minimum wage was announced in advance, some employers may have adjusted wages even before May 2011, potentially attenuating the observed before-after difference.

Finally, the use of midpoint approximations for earnings bands and setting the weeks of each month to be 4.345 introduces some measurement error in hourly wage estimates, yes remember:

(hours_main_employment or hours_7days) * 4.345 is monthly working hour

Then we are doing monthly MIDPOINT EARNINGS / monthly working hour to get the estimated hourly wage, then this


# Sidenote

Statistical significance is irrelevant. The question is whether the magnitude of the change is economically meaningful — and a drop from 23.5% to 12.9% clearly is.

This is actually more rigorous than tacking on p-values. Many economists (especially in labor economics) argue that with large administrative or survey data, significance testing is misleading because everything is significant.



# Robustness check (discarded)

## 1. Sample Composition Check (Addresses "Did low-skill workers disappear?")

```{r}
# Check if sample size changed for vulnerable groups
sample_check <- df_focus %>%
  filter(educ_group %in% c("No Schooling / Primary", "Lower Secondary"),
         year %in% c(2008:2012)) %>%
  group_by(year, educ_group) %>%
  summarise(
    weighted_n = sum(weight, na.rm = TRUE),
    raw_n = n(),
    .groups = "drop"
  )

print(sample_check)
```

Weighted N for No Schooling/Primary: 1.46M (2008) → 1.30M (2011) = -11% decline

Weighted N for Lower Secondary: 2.00M → 1.89M = -5.5% decline

Some low-skill workers left the sample (possible)

But decline is modest (not massive)

Could explain part of the proportion drop (Possible)

But not all — drop was 10+ percentage points



What to look for: If weighted_n drops sharply in 2011-2012, composition bias is possible. If it's stable, you can argue against it.



## 2. Upper Bound Robustness Test (Addresses measurement error)


```{r}
# Test a threshold that shouldn't be affected by minimum wage
threshold_check <- df_focus %>%
  filter(year >= 2008, year <= 2012) %>%
  group_by(year) %>%
  summarise(
    below_28 = weighted.mean(real_hourly_wage < 28, w = weight),
    below_80 = weighted.mean(real_hourly_wage < 80, w = weight),  # Robustness threshold
    .groups = "drop"
  )

print(threshold_check)

# If below_80 also dropped sharply in 2011, it's broad economic trend, not policy
```

Below $28: 8.77% → 5.93% = -2.84pp drop

Below $80: 63.5% → 63.4% = essentially flat (then drops to 59.6% in 2012)




## 3. High-Education Control Group (Addresses "was it just the economy?")

```{r}
# Compare low vs high education
control_check <- df_focus %>%
  filter(year %in% c(2008:2012),
         educ_broad %in% c("Lower Education", "Tertiary (Degree)")) %>%
  group_by(year, educ_broad) %>%
  summarise(
    below_28 = weighted.mean(real_hourly_wage < 28, w = weight) * 100,
    .groups = "drop"
  )

print(control_check)

# If both groups dropped similarly → macroeconomic trend
# If only low-education dropped → policy effect (but your Robustness says otherwise)
```

## Sample Composition Bias (Did low-skill workers disappear?)
第一点，就业那个点，我觉得很有必要验证一下
```{r}
# Check employment/population ratio for low-education groups over time
employment_check <- df_focus %>%
  filter(educ_group %in% c("No Schooling / Primary", "Lower Secondary"),
         year >= 2005, year <= 2012) %>%
  group_by(year, educ_group) %>%
  summarise(
    weighted_count = sum(weight, na.rm = TRUE),  # Estimated employed population
    sample_n = n(),
    .groups = "drop"
  ) %>%
  group_by(educ_group) %>%
  mutate(
    index_2008 = weighted_count / weighted_count[year == 2008] * 100  # 2008 = 100
  )

print(employment_check)

# Also check if this tracks with official HK employment statistics
# You can cite: C&SD Table 210-06101 (Employment by industry)
```

Key observations:

No Schooling/Primary dropped ~10-13% from 2008 -> Some composition change occurred

But most of the drop happened BEFORE 2011 (2008→2010: -13.5%) -> Not a 2011-specific phenomenon

2010→2011: 86.5 → 89.3 (actually increased slightly) -> No evidence of mass layoffs in 2011

Lower Secondary remained stable throughout (95-100) -> No composition bias for this group
以 2008 年为基准（=100），观察 2005-2012 年间样本人数是否发生明显变化。

「你看到的低薪比例下降，是因为雇主给低教育工人涨了工资，还是因为雇主直接把低效工人解雇了，导致他们在样本中消失了？」

结果显示了什么？

no schooling/primary 2008 to 2010 -> -13.5%, 2008 to 2011 -> -10.7%, 2011 年人数并未大幅下降，反而比 2010 年略回升

lower secondary, 2008 to 2010 -> -4.9%, 2008 to 2011 -> -5.5%, 人数基本稳定

人数下降主要发生在 2008-2010 年（最低工资实施之前）

2010→2011 年，低教育组人数没有继续下降（反而略有回升）

低薪比例下降幅度（-10.6 个百分点）远远大于样本人数下降幅度（-11%）

「样本构成变化不足以解释低薪比例的下降。而且人数下降主要发生在政策之前，政策实施后人数反而趋稳，说明下降主要来自‘同组内工资改善’，而非‘低薪工人被踢出样本』。」

```{r}
# Create the composition trend plot with data labels
employment_check %>%
  ggplot(aes(x = year, y = index_2008, color = educ_group, shape = educ_group)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_text(aes(label = round(index_2008, 1)), 
            vjust = -0.8, 
            size = 3, 
            show.legend = FALSE) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2011, linetype = "dotted", color = "red", alpha = 0.7) +
  annotate("text", x = 2011, y = 105, label = "SMW Implemented", 
           color = "red", size = 3.5, hjust = -0.1) +
  scale_x_continuous(breaks = seq(2005, 2012, 1)) +
  scale_y_continuous(limits = c(80, 110),
                     breaks = seq(80, 110, 5)) +
  labs(title = "Sample Composition of Low-Education Groups (2008 = 100)",
       subtitle = "Weighted employment counts indexed to 2008 baseline",
       x = "Year", 
       y = "Index (2008 = 100)",
       color = "Education Group",
       shape = "Education Group") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```


```{r}
employment_check %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = year, y = index_2008, fill = educ_group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  geom_text(aes(label = round(index_2008, 1)), 
            position = position_dodge(width = 0.7), 
            vjust = -0.5, 
            size = 3) +
  geom_hline(yintercept = 100, linetype = "dashed") +
  labs(title = "Sample Composition of Low-Education Groups (2008 = 100)",
       subtitle = "Weighted employment counts indexed to 2008 baseline",
       x = "Year", 
       y = "Index (2008 = 100)",
       fill = "Education Group") +
  scale_fill_manual(values = c("Lower Secondary" = "steelblue", 
                                "No Schooling / Primary" = "darkred")) +
  scale_y_continuous(limits = c(0, 120)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```














## Midpoint Error (Upper/Lower Bound Test)
```{r}
# Define upper and lower bounds for each earnings code
# Using your existing earnings_map structure

# Lower bound (use the lower end of each earnings band)
lower_bound_map <- tibble(
  code = str_pad(1:40, width = 2, pad = "0"),
  lower_hkd = c(0, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000,
                11000, 12000, 13000, 14000, 15000, 16000, 17000, 18000, 19000, 20000,
                22000, 24000, 26000, 28000, 30000, 32000, 34000, 36000, 38000, 40000,
                42000, 44000, 46000, 48000, 50000, 60000, 70000, 80000, 90000, 100000)
)

# Upper bound (use the upper end of each earnings band)
upper_bound_map <- tibble(
  code = str_pad(1:40, width = 2, pad = "0"),
  upper_hkd = c(1999, 2999, 3999, 4999, 5999, 6999, 7999, 8999, 9999, 10999,
                11999, 12999, 13999, 14999, 15999, 16999, 17999, 18999, 19999, 21999,
                23999, 25999, 27999, 29999, 31999, 33999, 35999, 37999, 39999, 41999,
                43999, 45999, 47999, 49999, 54999, 64999, 74999, 84999, 94999, 149999)
)

# Calculate hourly wages using upper and lower bounds
df_sensitivity <- df_focus %>%
  mutate(
    # Lower bound hourly wage
    earnings_main_lower = lower_bound_map$lower_hkd[match(earnings_main, lower_bound_map$code)],
    hourly_wage_lower = earnings_main_lower / (hours_main_employment * 4.345),
    
    # Upper bound hourly wage
    earnings_main_upper = upper_bound_map$upper_hkd[match(earnings_main, upper_bound_map$code)],
    hourly_wage_upper = earnings_main_upper / (hours_main_employment * 4.345),
    
    # Midpoint (your original)
    earnings_main_mid = earnings_map$midpoint_hkd[match(earnings_main, earnings_map$code)],
    hourly_wage_mid = earnings_main_mid / (hours_main_employment * 4.345)
  )

# Compare results across three methods for No Schooling group
sensitivity_results <- df_sensitivity %>%
  filter(educ_group == "No Schooling / Primary", year %in% c(2008:2012)) %>%
  group_by(year) %>%
  summarise(
    midpoint = weighted.mean(hourly_wage_mid < 28, w = weight, na.rm = TRUE) * 100,
    lower_bound = weighted.mean(hourly_wage_lower < 28, w = weight, na.rm = TRUE) * 100,
    upper_bound = weighted.mean(hourly_wage_upper < 28, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  )

print(sensitivity_results)
```

用三种不同的方法重新计算每小时工资，然后重新估算「无 schooling/小学」组中低于 HK$28 的比例：

中点值（Midpoint）,	原来用的方法（取月薪区间的中间值）
下限（Lower Bound), 取月薪区间的最低值（例如 code 01 = HK$0）
上限（Upper Bound）,	取月薪区间的最高值（例如 code 01 = HK$1,999）

「你的时薪是根据月薪层级中点值和 4.345 倍率算出来的，这本身带有巨大的测量误差。」

结果显示了什么？


关键发现：

    三种方法都显示 2010→2011 年出现大幅下降

    中点值：28.6% → 17.9%（-10.7 个百分点）

    下限：34.2% → 24.0%（-10.2 个百分点）

    上限：19.5% → 10.6%（-8.9 个百分点）

    中点和下限的下降幅度几乎相同，上限略小但方向一致


「无论我们假设月薪是区间的最低值、中间值还是最高值，2011 年的低薪比例都出现了显著下降。结论不依赖于中点值的选择，结果是稳健的。」





```{r}
# Prepare data for plotting
sensitivity_plot_data <- sensitivity_results %>%
  pivot_longer(cols = c(midpoint, lower_bound, upper_bound),
               names_to = "method",
               values_to = "proportion") %>%
  mutate(method = factor(method, 
                         levels = c("upper_bound", "midpoint", "lower_bound"),
                         labels = c("Upper Bound", "Midpoint", "Lower Bound")))

# Create the plot with data labels
ggplot(sensitivity_plot_data, aes(x = year, y = proportion, color = method, shape = method)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_text(aes(label = round(proportion, 1)), 
            vjust = -0.8, 
            size = 3.5, 
            show.legend = FALSE) +
  geom_vline(xintercept = 2011, linetype = "dashed", color = "red", alpha = 0.7) +
  annotate("text", x = 2011, y = 42, label = "SMW Implemented", 
           color = "red", size = 4, hjust = -0.1) +
  scale_x_continuous(breaks = seq(2008, 2012, 1)) +
  scale_y_continuous(limits = c(0, 45),
                     breaks = seq(0, 45, 5)) +
  labs(title = "Sensitivity Test: Proportion Below HK$28 by Earnings Assumption",
       subtitle = "No Schooling / Primary Workers | Three methods of calculating hourly wage",
       x = "Year", 
       y = "Proportion Below HK$28 (%)",
       color = "Assumption",
       shape = "Assumption") +
  scale_color_manual(values = c("Lower Bound" = "darkred",
                                 "Midpoint" = "steelblue",
                                 "Upper Bound" = "darkgreen")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10))
```








All three methods show a large decline in 2011 => Results are NOT sensitive to midpoint choice

Midpoint: 28.6% → 17.9% (-10.7pp) => main result

Lower bound: 34.2% → 24.0% (-10.2pp) => Similar magnitude 

Upper bound: 19.5% → 10.6% (-8.9pp) => Slightly smaller but still large

The spread between bounds narrowed in 2011-2012 => Suggests less measurement error at lower wages

```{r}
# ============================================================
# STEP 1: Create age_group in df_focus (if not already there)
# ============================================================

df_focus <- df_focus %>%
  mutate(
    age_num = as.numeric(age),
    age_group = case_when(
      age_num %in% 4:6   ~ "15-29 (Young)",
      age_num %in% 7:10  ~ "30-49 (Prime-age)",
      age_num %in% 11:14 ~ "50+ (Older)",
      TRUE               ~ "Unknown"
    )
  )

# ============================================================
# STEP 2: Calculate Education Group Composition
# ============================================================

edu_composition <- df_focus %>%
  filter(educ_group %in% c("No Schooling / Primary", "Lower Secondary"),
         year >= 2005, year <= 2012) %>%
  group_by(year, educ_group) %>%
  summarise(
    weighted_count = sum(weight, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(educ_group) %>%
  mutate(
    index_2008 = weighted_count / weighted_count[year == 2008] * 100,
    group_type = "Education Group"
  ) %>%
  rename(subgroup = educ_group) %>%
  select(year, subgroup, index_2008, group_type)

# ============================================================
# STEP 3: Calculate Age Group Composition
# ============================================================

age_composition <- df_focus %>%
  filter(age_group %in% c("15-29 (Young)", "50+ (Older)"),
         year >= 2005, year <= 2012) %>%
  group_by(year, age_group) %>%
  summarise(
    weighted_count = sum(weight, na.rm = TRUE),
    .groups = "drop"
  )

# Get 2008 baseline for age groups
baseline_2008_age <- age_composition %>%
  filter(year == 2008) %>%
  select(age_group, base_2008 = weighted_count)

# Calculate index
age_composition <- age_composition %>%
  left_join(baseline_2008_age, by = "age_group") %>%
  mutate(
    index_2008 = weighted_count / base_2008 * 100,
    group_type = "Age Group"
  ) %>%
  rename(subgroup = age_group) %>%
  select(year, subgroup, index_2008, group_type)

# ============================================================
# STEP 4: Combine Both Datasets
# ============================================================

composition_combined <- bind_rows(edu_composition, age_composition)

# ============================================================
# STEP 5: Check 50+ data is complete
# ============================================================

composition_combined %>% filter(subgroup == "50+ (Older)")

# ============================================================
# STEP 6: Create the Plot
# ============================================================

library(ggrepel)  # Optional: for better label placement

ggplot(composition_combined, aes(x = year, y = index_2008, 
                                  color = subgroup, shape = subgroup)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_text_repel(aes(label = round(index_2008, 0)), 
                  size = 3.0,
                  box.padding = 0.3,
                  point.padding = 0.5,
                  segment.color = NA,
                  show.legend = FALSE) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2011, linetype = "dotted", color = "red", linewidth = 0.8) +
  annotate("text", x = 2011, y = 135, label = "SMW Implemented (May 2011)", 
           color = "red", size = 3.5, hjust = -0.05) +
  scale_x_continuous(breaks = seq(2005, 2012, 1)) +
  scale_y_continuous(limits = c(70, 140),
                     breaks = seq(70, 140, 10)) +
  labs(title = "Sample Composition of Vulnerable Groups (2008 = 100)",
       subtitle = "Weighted employment counts indexed to 2008 baseline",
       x = "Year", 
       y = "Index (2008 = 100)",
       color = "Subgroup",
       shape = "Subgroup") +
  scale_color_manual(values = c("Lower Secondary" = "steelblue",
                                 "No Schooling / Primary" = "darkred",
                                 "15-29 (Young)" = "forestgreen",
                                 "50+ (Older)" = "purple")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


No Schooling / Primary = 下降在政策前

Lower Secondary = 基本稳定

15-29 (Young) = 先降后升，2012回到原点

50+ (Older) = 大幅增长

教授的质疑是：「你看到的低薪比例下降，是因为低薪工人涨了工资，还是因为他们被解雇了，从样本里消失了？」

低教育组的人数确实减少了（约 10%）

但减少主要发生在 2008-2010 年（政策前），2011 年后就稳定了

老年人的人数不但没减少，反而增加了 17-30%

「低薪工人被解雇了」? 低教育组人数减少，但发生在政策前 -> 所以不能用「政策导致人消失」来解释低薪比例的下降

「2011 年人数暴跌」? 2011 年后人数稳定，没有继续跌 -> 政策没有导致「被踢出去」

「老年工人也被解雇了」? 老年人人数增长了 30% -> 恰恰相反，更多老年人留在/进入劳动力市场


主要发现是：2011 年后低薪比例大幅下降。

因此：「人被踢出去了」这个替代解释被排除。你的主要发现（低薪比例下降）更可能来自工资上涨，而不是样本流失。


```{r fig.width=10, fig.height=10, out.width="100%"}
# ============================================================
# 修复后的图：使用 ggrepel 避免标签重叠
# ============================================================

library(ggrepel)

# 合并数据（确保使用最新的 age_composition）
composition_combined <- bind_rows(edu_composition, age_composition)

# 画图（使用 geom_text_repel 代替 geom_text）
ggplot(composition_combined, aes(x = year, y = index_2008, 
                                  color = subgroup, shape = subgroup)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  # 关键修改：用 geom_text_repel 替代 geom_text
  geom_text_repel(aes(label = round(index_2008, 0)), 
                  size = 3.0,
                  box.padding = 0.3,
                  point.padding = 0.5,
                  segment.color = NA,  # 不画连线（可选）
                  show.legend = FALSE) +
  geom_hline(yintercept = 100, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = 2011, linetype = "dotted", color = "red", linewidth = 0.8) +
  annotate("text", x = 2011, y = 135, label = "SMW Implemented \n(May 2011)", 
           color = "red", size = 3.5, hjust = -0.05) +
  scale_x_continuous(breaks = seq(2005, 2012, 1)) +
  scale_y_continuous(limits = c(70, 140),
                     breaks = seq(70, 140, 10)) +
  labs(title = "Sample Composition of Vulnerable Groups (2008 = 100)",
       subtitle = "Weighted employment counts indexed to 2008 baseline",
       x = "Year", 
       y = "Index (2008 = 100)",
       color = "Subgroup",
       shape = "Subgroup") +
  scale_color_manual(values = c("Lower Secondary" = "steelblue",
                                 "No Schooling / Primary" = "darkred",
                                 "15-29 (Young)" = "forestgreen",
                                 "50+ (Older)" = "purple")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 9),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```







```{r}
# ============================================================
# Sensitivity Test for Age Group (50+ Older Workers)
# Compare midpoint, lower bound, and upper bound estimates
# ============================================================

# Calculate hourly wages using upper and lower bounds (if not already done)
df_sensitivity <- df_focus %>%
  mutate(
    # Lower bound hourly wage
    earnings_main_lower = lower_bound_map$lower_hkd[match(earnings_main, lower_bound_map$code)],
    hourly_wage_lower = earnings_main_lower / (hours_main_employment * 4.345),
    
    # Upper bound hourly wage
    earnings_main_upper = upper_bound_map$upper_hkd[match(earnings_main, upper_bound_map$code)],
    hourly_wage_upper = earnings_main_upper / (hours_main_employment * 4.345),
    
    # Midpoint (your original)
    earnings_main_mid = earnings_map$midpoint_hkd[match(earnings_main, earnings_map$code)],
    hourly_wage_mid = earnings_main_mid / (hours_main_employment * 4.345)
  )

# Compare results across three methods for 50+ Older workers
sensitivity_results_age <- df_sensitivity %>%
  filter(age_group == "50+ (Older)", year %in% c(2008:2012)) %>%
  group_by(year) %>%
  summarise(
    midpoint = weighted.mean(hourly_wage_mid < 28, w = weight, na.rm = TRUE) * 100,
    lower_bound = weighted.mean(hourly_wage_lower < 28, w = weight, na.rm = TRUE) * 100,
    upper_bound = weighted.mean(hourly_wage_upper < 28, w = weight, na.rm = TRUE) * 100,
    .groups = "drop"
  )

print(sensitivity_results_age)

# Prepare data for plotting
sensitivity_plot_data_age <- sensitivity_results_age %>%
  pivot_longer(cols = c(midpoint, lower_bound, upper_bound),
               names_to = "method",
               values_to = "proportion") %>%
  mutate(method = factor(method, 
                         levels = c("upper_bound", "midpoint", "lower_bound"),
                         labels = c("Upper Bound", "Midpoint", "Lower Bound")))

# Create the plot with data labels
ggplot(sensitivity_plot_data_age, aes(x = year, y = proportion, color = method, shape = method)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_text(aes(label = round(proportion, 1)), 
            vjust = -0.8, 
            size = 3.5, 
            show.legend = FALSE) +
  geom_vline(xintercept = 2011, linetype = "dashed", color = "red", alpha = 0.7) +
  annotate("text", x = 2011, y = 35, label = "SMW Implemented", 
           color = "red", size = 4, hjust = -0.1) +
  scale_x_continuous(breaks = seq(2008, 2012, 1)) +
  scale_y_continuous(limits = c(0, 45),
                     breaks = seq(0, 45, 5)) +
  labs(title = "Sensitivity Test: Proportion Below HK$28 by Earnings Assumption",
       subtitle = "Older Workers (50+) | Three methods of calculating hourly wage",
       x = "Year", 
       y = "Proportion Below HK$28 (%)",
       color = "Assumption",
       shape = "Assumption") +
  scale_color_manual(values = c("Lower Bound" = "darkred",
                                 "Midpoint" = "steelblue",
                                 "Upper Bound" = "darkgreen")) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10))


```



```{r}
# Threshold test for HK$23, HK$28, HK$33
threshold_test <- df_focus %>%
  filter(educ_group == "No Schooling / Primary", year %in% seq(2000, 2012)) %>%
  group_by(year) %>%
  summarise(
    below_23 = weighted.mean(real_hourly_wage < 23, w = weight, na.rm = TRUE) * 100,
    below_28 = weighted.mean(real_hourly_wage < 28, w = weight, na.rm = TRUE) * 100,
    below_33 = weighted.mean(real_hourly_wage < 33, w = weight, na.rm = TRUE) * 100
  )

print(threshold_test)

# Prepare data for plotting
threshold_plot_data <- threshold_test %>%
  pivot_longer(cols = c(below_23, below_28, below_33),
               names_to = "threshold",
               values_to = "proportion") %>%
  mutate(threshold = factor(threshold,
                            levels = c("below_23", "below_28", "below_33"),
                            labels = c("Below HK$23", "Below HK$28", "Below HK$33")))

# Create the plot
ggplot(threshold_plot_data, aes(x = year, y = proportion, color = threshold, shape = threshold)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  geom_text(aes(label = round(proportion, 1)),
            vjust = -0.8,
            size = 3,
            show.legend = FALSE) +
  geom_vline(xintercept = 2011, linetype = "dashed", color = "red", alpha = 0.7) +
  annotate("text", x = 2011, y = 35, label = "SMW",
           color = "red", size = 3.5, hjust = -0.05) +
  scale_x_continuous(breaks = seq(2000, 2012, 2)) +
  #scale_y_continuous(limits = c(0, 35),                     breaks = seq(0, 35, 5)) +
  labs(title = "Proportion Below Different Thresholds (HK$23, HK$28, HK$33)",
       subtitle = "No Schooling / Primary Workers | Real wages adjusted to 2011 Q2 prices",
       x = "Year",
       y = "Proportion (%)",
       color = "Threshold",
       shape = "Threshold") +
  scale_color_manual(values = c("Below HK$33" = "darkgreen",
                                 "Below HK$28" = "steelblue",
                                 "Below HK$23" = "darkred")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 10))

```










